---
title: "Ryan Tillis - Data Science - R Programming - Quiz 1 - Coursera"
author: <a href="http://www.ryantillis.com"> Ryan Tillis </a>
date: "May 24, 2016"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quiz 1

This is Quiz 1 from Coursera's R Programming class within the <a href="https://www.coursera.org/specializations/jhu-data-science">Data Science Specialization</a>. This publication is intended as a learning resource, all answers are documented and explained. Datasets are available in R packages.

<hr>

<font size="+2">**1.**  </font>The R language is a dialect of which of the following programming languages?

<hr>

* <font size="+1">**S**</font>

<hr>

#####Explanation:

R is an open source implementation of S with a revised syntax and an awesome community.

```{r Question 1}

```

<hr>
<font size="+2">**2.**  </font> The definition of free software consists of four freedoms (freedoms 0 through 3). Which of the following is NOT one of the freedoms that are part of the definition? Select all that apply.

<hr>

<font size="+1"><b>

* The freedom to sell the software for any price.

* The freedom to restrict access to the source code for the software.

* The freedom to prevent users from using the software for undesirable purposes.

</b>
</font>

<hr>

#####Explanation:

Yay free software!

```{r Q2}
dat <- read.table('http://www4.stat.ncsu.edu/~stefanski/NSF_Supported/Hidden_Images/orly_owl_files/orly_owl_Lin_4p_5_flat.txt', header = FALSE)
fit <- lm(V1 ~ . - 1, data = dat); plot(predict(fit), resid(fit), pch = '.')
```
<hr>

<font size="+2">3.  </font> In R the following are all atomic data types EXCEPT: (Select all that apply)

<hr>

<font size="+1">
<b>

* list

* array

* matrix

* data frame

* table

</b>
</font>

<hr>

#####Explanation:

Predicting with the lower and upper bounds of the confidence intervals

```{r Question 3}
fit <- lm(mpg~wt,mtcars)
summary(fit)$coef
```

<hr>

<font size="+2">**4.**  </font> If I execute the expression x <- 4 in R, what is the class of the object `x' as determined by the `class()' function?

<hr>

* <font size="+1">**numeric
**</font>

<hr>

#####Explanation:

R automatically interprets 4 as a numeric class object.

```{r Question 4}
```

<hr>

<font size="+2">5.  </font>What is the class of the object defined by the expression x <- c(4, "a", TRUE)?

<hr>

* <font size="+1">**character**</font>

<hr>

#####Explanation:



```{r Question 5}
x <- c(4, "a", TRUE)
class(x)
```

<hr>

<font size="+2">**6.**  </font> If I have two vectors x <- c(1,3, 5) and y <- c(3, 2, 10), what is produced by the expression cbind(x, y)?
<hr>

* <font size="+1">**a matrix with 2 columns and 3 rows**</font>

<hr>

#####Explanation:

Combine the two vectors as columns.

```{r Question 6}
```

<hr>

<font size="+2">**7.**  </font> A key property of vectors in R is that

<hr>

* <font size="+1">**elements of a vector all must be of the same class**</font>

<hr>

#####Explanation:

This is nice for statistical purposes.

```{r Question 7}
```

<hr>

<font size="+2">**8.**  </font> Suppose I have a list defined as x <- list(2, "a", "b", TRUE). What does x[[2]] give me? Select all that apply.

<hr>

<font size="+1"><b>

* a character vector containing the letter "a".

* a character vector of length 1.

</b>
</font>

<hr>

#####Explanation:

Two brackets gives the actual element inside of the list, one bracket gives the list with the element inside.

<hr>

<font size="+2">**9.**  </font> Suppose I have a vector x <- 1:4 and a vector y <- 2. What is produced by the expression x + y?

<hr>

* <font size="+1">**a numeric vector with elements 3, 4, 5, 6.**</font>

<hr>

#####Explanation:



```{r Question 9}

```

<hr>

<font size="+2">**10.**  </font> Suppose I have a vector x <- c(17, 14, 4, 5, 13, 12, 10) and I want to set all elements of this vector that are greater than 10 to be equal to 4. What R code achieves this? Select all that apply.

<hr>

<font size="+1"><b>

* x[x >= 11] <- 4

* x[x > 10] <- 4

</b>
</font>

<hr>

#####Explanation:

Indexing with a boolean.

```{r Question 10}

```

<hr>

<font size="+2">**11.**  </font>Use the Week 1 Quiz Data Set to answer questions 11-20.

In the dataset provided for this Quiz, what are the column names of the dataset?
<hr>

* <font size="+1">**Ozone, Solar.R, Wind, Temp, Month, Day**</font>

<hr>

#####Explanation:

Download, unzip, read, print.

```{r Question 11}
dat <- download.file('https://d396qusza40orc.cloudfront.net/rprog/data/quiz1_data.zip', destfile ="quizdat.zip")
dat <- unzip("quizdat.zip")
dat <- read.csv("hw1_data.csv")
names(dat)
```

<hr>
<font size="+2">**12.**  </font> Extract the first 2 rows of the data frame and print them to the console. What does the output look like?

<hr>

#####Explanation:

Index

```{r Q12}
dat[1:2,]
```
<hr>

<font size="+2">13.</font> How many observations (i.e. rows) are in this data frame?

<hr>

* <font size="+1">**153**</font>

<hr>

#####Explanation:

Nrow()

```{r Question 13}
nrow(dat)
```

<hr>

<font size="+2">**14.**  </font> Extract the last 2 rows of the data frame and print them to the console. What does the output look like?

<hr>

#####Explanation:

Correlation(XY)* SDy/SDx

```{r Question 14}
dat[152:153,]
```

<hr>

<font size="+2">15.</font>What is the value of Ozone in the 47th row?

<hr>

* <font size="+1">**21**</font>

<hr>

#####Explanation:

$ notation is useful

```{r Question 15}
dat$Ozone[47]
```

<hr>

<font size="+2">**16.**  </font> How many missing values are in the Ozone column of this data frame?
<hr>

* <font size="+1">**37**</font>

<hr>

#####Explanation:

Is NA return T/F values which can be summed to get a count of NAs.

```{r Question 16}
sum(is.na(dat$Ozone))
```

<hr>

<font size="+2">**17.**  </font> What is the mean of the Ozone column in this dataset? Exclude missing values (coded as NA) from this calculation.
<hr>

* <font size="+1">**42.1**</font>

<hr>

#####Explanation:

na.rm is a great option for calculation where NAs might interfere

```{r Question 17}
mean(dat$Ozone, na.rm=TRUE)
```

<hr>

<font size="+2">**18.**  </font> Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90. What is the mean of Solar.R in this subset?

<hr>

* <font size="+1">**212.8**</font>

<hr>

#####Explanation:

Which give index of booleans, $ selects columns.

```{r Question 18}
mean(dat[which(dat$Ozone >31 & dat$Temp > 90),]$Solar.R)
```

<hr>

<font size="+2">**19.**  </font> What is the mean of "Temp" when "Month" is equal to 6?

<hr>

* <font size="+1">**79.1**</font>

<hr>

#####Explanation:

Same as above

```{r Question 19}
mean(dat[which(dat$Month == 6),]$Temp)
```

<hr>

<font size="+2">**20.**  </font> Let the slope having fit Y as the outcome and X as the predictor be denoted as β1. Let the slope from fitting X as the outcome and Y as the predictor be denoted as γ1. Suppose that you divide β1 by γ1; in other words consider β1/γ1. What is this ratio always equal to?

<hr>

* <font size="+1">**115**</font>

<hr>

#####Explanation:

Need to remove NA for this.

```{r Question 20}
max(dat[which(dat$Month == 5),]$Ozone, na.rm = TRUE)
```

<hr>
Check out my website at: <http://www.ryantillis.com/>
<hr>
